---
title: "Complete Rio de Janeiro Downloads"
author: "Ryan Santos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Complete Rio de Janeiro Downloads}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why Completeness Matters

For Rio de Janeiro, the package expects 92 municipalities. A table filtered to
RJ is not automatically complete. Completeness must be measured against the
internal registry and, when time columns exist, by year or by year-month.

The executable examples below use simulated data so this vignette builds
offline.

```{r}
library(vigiar)

rj <- vigiar_rj_municipios()
code_col <- if ("codigo_ibge_6" %in% names(rj)) "codigo_ibge_6" else "codigo_ibge"

coverage_fun <- function(dados, por = "geral") {
  if (exists("vigiar_rj_cobertura", mode = "function")) {
    return(vigiar_rj_cobertura(dados, por = por))
  }

  if (por == "ano" && "ano" %in% names(dados)) {
    years <- sort(unique(dados$ano))
    return(do.call(rbind, lapply(years, function(y) {
      out <- coverage_fun(dados[dados$ano == y, ], por = "geral")
      out$nivel <- "ano"
      out$ano <- y
      out
    })))
  }

  present <- unique(dados$cod_municipio)
  present <- present[present %in% rj[[code_col]]]
  data.frame(
    nivel = por,
    ano = NA_integer_,
    mes = NA_integer_,
    n_municipios_presentes = length(present),
    n_municipios_esperados = nrow(rj),
    cobertura_pct = round(100 * length(present) / nrow(rj), 1),
    n_ausentes = nrow(rj) - length(present),
    completo = length(present) == nrow(rj),
    stringsAsFactors = FALSE
  )
}

missing_fun <- function(dados) {
  if (exists("vigiar_rj_municipios_ausentes", mode = "function")) {
    return(vigiar_rj_municipios_ausentes(dados))
  }
  present <- unique(dados$cod_municipio)
  rj[!rj[[code_col]] %in% present, ]
}

head(rj)
```

## Simulated Complete Coverage

```{r}
pm25_complete <- data.frame(
  cod_municipio = rj[[code_col]],
  ano = 2022L,
  pm25_media_anual = 12,
  stringsAsFactors = FALSE
)

coverage_fun(pm25_complete)
```

## Simulated Partial Coverage

```{r}
pm25_partial <- pm25_complete[1:10, ]

cov_partial <- coverage_fun(pm25_partial)
cov_partial

missing_fun(pm25_partial)
```

## Coverage by Year

```{r}
pm25_years <- rbind(
  transform(pm25_complete, ano = 2020L),
  transform(pm25_partial, ano = 2021L)
)

coverage_fun(pm25_years, por = "ano")
```

## Table-Aware Completeness

`vigiar_rj_completude_tabela()` applies the expected municipal grid for common
VIGIAR tables. For `df_anual`, the expected grid is municipality by year. For
`df_mensal`, the expected grid is municipality by year by month.

```{r}
if (exists("vigiar_rj_completude_tabela", mode = "function")) {
  vigiar_rj_completude_tabela(pm25_years, tabela = "df_anual")
}
```

When `require_complete = TRUE`, incomplete coverage or possible truncation is an
error, not just a warning.

```{r, eval = FALSE}
vigiar_rj_completude_tabela(
  pm25_years,
  tabela = "df_anual",
  require_complete = TRUE
)
```

## Municipality-Specific Download

Use `vigiar_baixar_municipio()` when the target is one municipality. It filters
by normalized IBGE code, not by municipality name.

```{r, eval = FALSE}
campos <- vigiar_baixar_municipio("df_anual", codigo_ibge = 330100)

attr(campos, "vigiar_codigo_ibge_6")
attr(campos, "vigiar_codigo_ibge_7")
attr(campos, "vigiar_municipio")
```

## Online Download Example

The following code is intentionally not evaluated during vignette builds.

```{r, eval = FALSE}
vigiar_conectar()

pm25_rj <- vigiar_baixar_rj("df_anual", validar_cobertura = TRUE)
vigiar_rj_cobertura(pm25_rj)

pm25_rj_complete <- vigiar_baixar_rj(
  "df_anual",
  validar_cobertura = TRUE,
  require_complete = TRUE
)

vigiar_desconectar()
```

## Possible Truncation

Large Power BI responses can be truncated. `vigiar_baixar_rj()` attaches the
`vigiar_possivel_truncamento` attribute when the response size reaches a known
or requested limit. Do not claim 92/92 completeness from a truncated response
without a validated partitioned download.

```{r}
attr(pm25_partial, "vigiar_possivel_truncamento") <- TRUE
coverage_fun(pm25_partial)
```

## Scientific Interpretation

Coverage tells you whether the expected municipalities appear in the downloaded
table. It does not prove that PM2.5 is measured at the ground, that the data are
fit for causal inference, or that the time series is sufficient for modelling.
The package does not implement GAM, DLNM, relative-risk, machine-learning, or
causal inference models. Treat coverage as a necessary diagnostic, not as a
scientific conclusion.
